---
title: "Recommender system Assignment"
output: html_document
---

It is my second attempt to finish the specialization 
on Recommender systems from [University of Minnesota](https://www.coursera.org/specializations/recommender-systems) but unlike the first time, I am attempting to complete all the assignments not only in Java but also in R - for the practice sake.

### Course #1 - Introduction to Recommender Systems: Non-Personalized and Content-Based

The first assignment in the course ["Introduction to Recommender Systems: Non-Personalized and Content-Based"](https://www.coursera.org/learn/recommender-systems-introduction) will "explore non-personalized and lightly-personalized recommendations" and includes sever deliverables with each deliverable representing a different analysis of the data provided."The first three deliverables represent non-personalized summary statistics; the next two represent product association using two different mechanisms (we aren't using lift here because of how similar the popularity levels are). The last two represent a demographic analysis to explore whether using gender would be wise given this dataset."

```{r }
## Read the dataset uploaded at Google Drive
install.packages('gsheet', repos="http://cran.rstudio.com/")
library('gsheet')
url <- 'https://docs.google.com/spreadsheets/d/1XDBRCYFTxsw27AivxJ5pWxDHN0WA6GqSP46PVe2BCQ4/edit?usp=sharing'
dataset <- gsheet2tbl(url)
```

1. Deliverable i.e. calulcation of mean rating. 

**Task**: calculate the mean rating for each movie, order with the highest rating listed first, and submit the top three (along with the mean scores for the top two).

```{r}

dataset_mean_1 <- data.frame(colMeans(dataset, na.rm = TRUE))
install.packages('plyr', repos="https://cran.r-project.org")
library('plyr')
renamed_mean_1 <- rename(dataset_mean_1,c('colMeans.dataset..na.rm...TRUE.'='Mean'))
ordered_mean_1 <- head(renamed_mean_1[order(-renamed_mean_1$Mean),,drop=FALSE],n=4)
```

The ranking of top 3 movies based on average.

| Movie                                          | Score |
|------------------------------------------------|-------|
| 318: Shawshank Redemption, The (1994)          | 3.6   |
| 260: Star Wars: Episode IV - A New Hope (1977) | 3.27  |
| 541: Blade Runner (1982)                       | 3.22  |

2. Deliverable i.e. rating count (popularity)

**Task**: Count the number of ratings for each movie, order with the most number of ratings first, and submit the top three (along with the counts for the top two).
```{r}
dataset_count <- data.frame(colSums(!is.na(dataset)))
library('plyr')
renamed_ordered_count <- rename(dataset_count, c('colSums..is.na.dataset..'='Mean'))
ordered_count <- head(renamed_ordered_count[order(-renamed_ordered_count$Mean),,drop=FALSE],n=5)
```

The ranking of top 3 movies based on count.

| Movie                                          | Count |
|------------------------------------------------|-------|
| 1: Toy Story (1995)                            | 17    |
| 593: Silence of the Lambs, The (1991)          | 16    |
| 260: Star Wars: Episode IV - A New Hope (1977) | 15    |
